---
title: "Assignment 8"
author: "GW"
date: "2025-03-11"
output: html_document
---

Loading in packages.

```{r, message = F}
library(ggplot2)
library(dplyr)
library(vegan)
library(ape)
library(ggtree)
```

Loading in data and inspecting.

```{r}
dat <- read.csv("C:/Users/Grace/Downloads/FloristicSurvey.csv")
```


head(dat)
tail(dat)
dim(dat)
summary(dat)
```

Selecting for data to be used in analysis. 

```{r}
spec_dat <- dat[, 10:ncol(dat)]
rownames(spec_dat) <- dat$Quadrate
```

Creating an OTU table. 

```{r}
otu_table <- as.data.frame(spec_dat)
head(otu_table)
```

NMDS

```{r, message = F}
set.seed(10)
nmds_dat <- metaMDS(otu_table, distance = "bray", 
                    binary = F)
```

Combining datasets.

```{r}
dat2 <- data.frame(NMDS1 = nmds_dat$points[,1],
                  NMDS2 = nmds_dat$points[,2],
                  Quadrate = row.names(otu_table))
```

```{r}
merge_dat <- merge(dat2, dat, by = "Quadrate", all.x = T, all.y = F)
```


Plotting NMDS with Location (in vs. out)

```{r}
ggplot(merge_dat, aes(x = NMDS1, y = NMDS2, colour = Location)) +
  geom_point() +
  theme_minimal()
```

```{r}
ggplot(merge_dat, aes(x = NMDS1, y = NMDS2, colour = as.factor(Population))) +
  geom_point() +
  theme_minimal()
```

Creating a tree.

```{r}
dist <- vegdist(spec_dat, method = "bray", binary = F)
spec_tree <- nj(dist)
ggtree(spec_tree, layout = "rectangular") %<+%
  dat + geom_tiplab(aes(colour = as.factor(Population))) +
  geom_label(aes(x = branch, label = Location), fill = 'lavender') 
```

